---
title: "Cytokines"
output: html_document
---

Makes the cytokine plots
Statistic use non-debatched data - Rfit rank-based linear model including cohort plus either strain or condition.
Visualization uses debatched data

Makes:
- Figure S3B
- Figure S9A

```{r}
library(ggplot2)
library(here)
library(tidyverse)
library(magrittr)

load(here("res/data_imp.R")) # Non-debatched data
load(here("res/apollo_together.RData")) # Debatched data

prop_above_ll <- 0.9

colnames(data_imp) %<>% gsub("\\+", "pos", .)
colnames(data_imp) %<>% gsub("\\-", "neg", .)

inv_pls <- inv_limma_pls 
deb <- limma_deb

star <- function(pval) {
  if (pval < 0.001) {
		return("***")
		}
	if (pval < 0.01) {
		return("**")
		}		
	if (pval < 0.05) {
		return("*")
	}
  #if (pval < 0.1) {
	#	return("+")
	#	}
	else {return("")
		}
		}

options(scipen=100, digits=6)

```


For cytokines, check conditions and strains were balanced across plates
```{r}
print("Is plate confounded with strain, condition or cohort?")
glm_fit <- glm(family=binomial, data=data_imp, Plate ~ strain + cond + factor(coh))
summary(glm_fit) # Not confounded
car::Anova(glm_fit, type="2")
```

Homecage CD19 vs. WT plots
```{r}
library(grid)

cytokine_labels <- feature_labels[names(feature_labels) %in% cytokines]

library(Rfit)
ps <- vector(length=length(cytokines))
ps_fits <- list()
for (i in 1:length(cytokines)){
  subset <- forpls[forpls$cond=="hc",c("strain","coh",cytokines[i])]
  colnames(subset)[3] <- "value"
  subset$coh <- droplevels(subset$coh)
  tmp <- rfit(value ~ strain + coh, data=subset) 
  print(summary(tmp)$coefficients)
  ps[i] <- summary(tmp)$coefficients["straincd19","p.value"]
  ps_fits[[i]] <- tmp
}
ps_selection_adj_rfit <- p.adjust(ps,method="BH")

print("P and adjusted P for effect of strain")
print(cbind(cytokines,signif(ps,3),signif(ps_selection_adj_rfit,3)))

# Check an example fit
rf <- ps_fits[[9]]
plot(fitted.values(rf), rstudent(rf)); abline(h = c(-2, 2)) 
qqnorm(residuals(rf)); qqline(residuals(rf)) 

```

Plot effects of CD19 deficiency on cytokine levels (in homecage animals)
```{r}

# Graph of boxplots
plotCytokines <- function(i=NULL, data=NULL, labs=NULL, p_raw=NULL, p_adj=NULL, show=NULL){
    name <- names(labs[i])
    label <- labs[i]
    subset <- data.frame(strain=data[data$cond=="hc","strain"], choice=data[data$cond=="hc",name])
    ifelse(is.null(p_adj), siggy <- paste("p=",signif(p_raw[i],1),sep=""), siggy <- paste("p(adj)=",signif(p_adj[i],1),sep=""))
    ifelse(is.null(p_adj), starry <- star(p_raw[i]), starry <- star(p_adj[i]))
    ifelse(is.null(show), showadd <- textGrob(siggy, x = 0.25, y = 0.9, just="centre"), showadd <- textGrob(starry, x = 0.5, y = 0.95, just="centre", gp=gpar(fontsize=20)))
    ggplot(data=subset, aes(x=strain, y=choice)) + geom_jitter(width=0.15, shape=16) + stat_summary(fun.y=median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5, colour="red") + 
    ggtitle(parse(text=label)) + 
    xlab("") + ylab("") + theme_bw() + theme(axis.text.x=element_blank()) +
    annotation_custom(showadd, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) 
}

# Graph of points + horizontal bar
plotCytokines2 <- function(i=NULL, data=NULL, labs=NULL, p_raw=NULL, p_adj=NULL, show=NULL, cols=NULL){
    name <- names(labs[i])
    label <- labs[i]
    subset <- data.frame(strain=data[data$cond=="hc","strain"], choice=data[data$cond=="hc",name])
    ylim1 = boxplot.stats(subset$choice)$stats[c(1, 5)] 
    ifelse(is.null(p_adj), siggy <- paste("p=",signif(p_raw[i],1),sep=""), siggy <- paste("p(adj)=",signif(p_adj[i],1),sep=""))
    ifelse(is.null(p_adj), starry <- star(p_raw[i]), starry <- star(p_adj[i]))
    ifelse(is.null(show), showadd <- textGrob(siggy, x = 0.25, y = 0.9, just="centre"), showadd <- textGrob(starry, x = 0.5, y = 0.95, just="centre", gp=gpar(fontsize=20)))
    ggplot(data=subset, aes(x=strain, y=choice, fill=strain)) + 
    ggtitle(parse(text=label)) + 
    xlab("") + ylab("") +
    geom_boxplot(outlier.shape=NA) + guides(fill=FALSE) + scale_fill_manual(values = cols) + ylim(limits=1.05*ylim1) +
    annotation_custom(showadd, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
      scale_x_discrete(labels=c(parse(text="WT"),parse(text="CD19^{'-/-'}"))) +
      theme_bw() + 
  theme(axis.title=element_text(size=20), axis.text.y=element_text(size=14), axis.text.x=element_blank(), plot.title=element_text(size=20))
}

# Test
plotCytokines(data=inv_pls, i=2, labs=cytokine_labels, p_raw=ps, p_adj=NULL)
plotCytokines2(data=inv_pls, i=2, labs=cytokine_labels, p_raw=ps, p_adj=NULL, cols=c("light grey","dark green"))

library(gridExtra)
library(cowplot)
p <- list(); p2 <- list(); p3 <- list(); p4 <- list()

# Make plots using the debatched data with stats from the Rfit models
for (i in 1:length(names(cytokine_labels))){
  p2[[i]] <- plotCytokines(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps)
  p3[[i]] <- plotCytokines2(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps, show="starry", cols=c("light grey","dark green"))
  p4[[i]] <- plotCytokines2(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps, cols=c("light grey","dark green"))
}

# Figures S9A
g <- do.call(grid.arrange,c(p3, ncol=length(names(cytokine_labels))))
save_plot(filename = paste("apollo_deb_cytokines_hc_prop_above_ll_padj_box",prop_above_ll,".pdf",sep=""), g, path=here("pics"), base_width=1.7*length(names(cytokine_labels)), base_height=3)

g <- do.call(grid.arrange,c(p4, ncol=length(names(cytokine_labels))))
save_plot(filename = paste("deb_cytokines_hc_prop_above_ll_padj_num_box",prop_above_ll,".pdf",sep=""), g, path=here("pics"), base_width=2.5*length(names(cytokine_labels)), base_height=4)

```

Stats and plots for HC vs. SD (in WT animals)
```{r}
ps <- vector(length=length(cytokines))
for (i in 1:length(cytokines)){
  subset <- forpls[forpls$strain=="wt",c("cond","coh",cytokines[i])]
  colnames(subset)[3] <- "value"
  subset$coh <- droplevels(subset$coh)
  tmp <- rfit(value ~ cond + coh, data=subset)
  print(summary(tmp)$coefficients)
  ps[i] <- summary(tmp)$coefficients["condcsd","p.value"]
}
ps_selection_adj_rfit <- p.adjust(ps,method="BH")
print("P and adjusted P for effect of condition")
print(cbind(cytokines,signif(ps,3),signif(ps_selection_adj_rfit,3)))

plotCytokines <- function(i=NULL, data=NULL, labs=NULL, p_raw=NULL, p_adj=NULL, show=NULL){
    name <- names(labs[i])
    label <- labs[i]
    subset <- data.frame(cond=data[data$strain=="wt","cond"], choice=data[data$strain=="wt",name])
    ifelse(is.null(p_adj), siggy <- paste("p=",signif(p_raw[i],1),sep=""), siggy <- paste("p(adj)=",signif(p_adj[i],1),sep=""))
    ifelse(is.null(p_adj), starry <- star(p_raw[i]), starry <- star(p_adj[i]))
    ifelse(is.null(show), showadd <- textGrob(siggy, x = 0.25, y = 0.9, just="centre"), showadd <- textGrob(starry, x = 0.5, y = 0.95, just="centre", gp=gpar(fontsize=20)))
    ggplot(data=subset, aes(x=cond, y=choice)) + geom_jitter(width=0.15, shape=16) + stat_summary(fun.y=median, fun.ymin = median, fun.ymax = median, geom = "crossbar", width = 0.5, colour="red") +
      theme_bw() + theme(axis.text.x=element_blank()) +
    ggtitle(parse(text=label)) + 
    xlab("") + ylab("") +
      annotation_custom(showadd, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf)
}

plotCytokines2 <- function(i=NULL, data=NULL, labs=NULL, p_raw=NULL, p_adj=NULL, show=NULL, cols=NULL){
    name <- names(labs[i])
    label <- labs[i]
    subset <- data.frame(cond=data[data$strain=="wt","cond"], choice=data[data$strain=="wt",name])
    ylim1 = boxplot.stats(subset$choice)$stats[c(1, 5)]
    ifelse(is.null(p_adj), siggy <- paste("p=",signif(p_raw[i],1),sep=""), siggy <- paste("p(adj)=",signif(p_adj[i],1),sep=""))
    ifelse(is.null(p_adj), starry <- star(p_raw[i]), starry <- star(p_adj[i]))
    ifelse(is.null(show), showadd <- textGrob(siggy, x = 0.25, y = 0.9, just="centre"), showadd <- textGrob(starry, x = 0.5, y = 0.95, just="centre", gp=gpar(fontsize=20)))
    ggplot(data=subset, aes(x=cond, y=choice, fill=cond)) + 
    ggtitle(parse(text=label)) + 
    geom_boxplot(outlier.shape=NA) + guides(fill=FALSE) + scale_fill_manual(values = cols) + ylim(limits=1.05*ylim1) +
    xlab("") + ylab("") +
    annotation_custom(showadd, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +
      scale_x_discrete(labels=c("HC",parse(text="Stress"))) +
  
      theme_bw() + theme(axis.title=element_text(size=20), axis.text.y=element_text(size=14), axis.text.x=element_blank(), plot.title=element_text(size=20))
}

library(gridExtra)
p <- list(); p2 <- list(); p3 <- list(); p4 <- list()
for (i in 1:length(cytokines)){
  p[[i]] <- plotCytokines(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps, show="starry") # Choosing adjusted
  p2[[i]] <- plotCytokines(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps) # Choosing adjusted with numbers
  p3[[i]] <- plotCytokines2(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps, show="starry", cols=c("light grey","red")) # Choosing adjusted
  p4[[i]] <- plotCytokines2(data=inv_pls, i=i, labs=cytokine_labels, p_adj=ps_selection_adj_rfit, p_raw=ps, cols=c("light grey","red")) # Choosing adjusted with numbers
}

g <- do.call(grid.arrange,c(p2, ncol=length(cytokines)))
save_plot(filename = paste("deb_cytokines_wt_prop_above_ll_padj_num", prop_above_ll,".pdf",sep=""), g, path=here("pics"), base_width=1.7*length(cytokines), base_height=3)

# Figure S3B
g <- do.call(grid.arrange,c(p3, ncol=length(cytokines)))
save_plot(filename = paste("apollo_deb_cytokines_wt_prop_above_ll_padj_box", prop_above_ll,".pdf",sep=""), g, path=here("pics"), base_width=1.7*length(cytokines), base_height=3)

```

```{r}
sessionInfo()
```
