
Import and QC for myeloid splenic data

```{r}
basedir='/Users/mary/data/nih/'
setwd('/Users/mary/data/nih/')
picpath=paste(basedir,"pics/",sep="")

fcs_path <- paste(basedir,"facs/",sep="")
flow.dir <- paste(basedir,"fcs_analysis",sep="")
flow.files <- list.files(path = flow.dir, pattern = "r[0-9]_my_v2.wsp$", all.files = FALSE, full.names = TRUE, recursive = TRUE, ignore.case = TRUE) 
print(flow.files)
```

This is for MYELOID SPLEEN
Make flowSet with just the cells I want in the analysis 
- including all cells in the "live" gate
- also need to include information on manual labelling, hence including gates
- have taken all mice together (wt, cd19, hc, stress)
- using flowJo compensation and transformations as it is

```{r}
load(paste(basedir,'data_bred.R',sep=""))
load(paste(basedir,"/analysis/r/counts_good_names_spl.Rmd",sep=""))
```

```{r}
# Define gates for manual gating
gateNames <- c("cdcs_cd11b","cdcs_cd103","cdcs_dn","cdcs_cd103cd11b","pDCs","neutrophils","f4_80_macs","inflamm_mono","Ly6Cpos_CD11b_neg") # must be in same order as counts_good_names_my 
counts_gate_names_my <- gateNames
counts_gate_names_my[5:9] <- counts_good_names_spl[10:14]
counts_good_names_my <- counts_good_names_spl[8:14]
cbind(gateNames, counts_gate_names_my) # Check equivalence

gateNamesPlus <- c(gateNames,"live") 

rm(fcs_raw_list) # The flowset 
rm(res_rawer) # The per-cell manual gating information

library(flowWorkspace)
for ( i in 1:length(flow.files) ){
  print(flow.files[i])
  ws <- openWorkspace(flow.files[i])
  print(ws)
  b <- parseWorkspace(ws, name="myeloid_spleen", path=fcs_path,cleanup=FALSE,keep.indices=TRUE,includeGates=TRUE, execute=TRUE)
  fcs <- getData(b) 
  res <- list()
  for (p in 1:length(fcs)){
    res[[p]] <- matrix(nrow=dim(getData(b[[p]])@exprs)[1],ncol=length(gateNamesPlus),dimnames=list(NULL,gateNamesPlus)) # Make empty indices matrix
  for (g in 1:length(gateNamesPlus)){
    res[[p]][,gateNamesPlus[g]] <- getIndices(b[[p]],gateNamesPlus[g])
  }
  }
  if (exists("res_rawer")) { res_rawer <- c(res_rawer,res) } else { res_rawer <- res }
  if (exists("fcs_raw_list")) { fcs_raw_list <- c(fcs_raw_list,fcs) } else { fcs_raw_list <- fcs } #Annoyingly, makes list but not same colnames so can't just rbind2
closeWorkspace(ws)
}

cbind(colnames(fcs_raw_list[[1]][[1]]@exprs), colnames(fcs_raw_list[[5]][[1]]@exprs))

library(flowCore)
tmp <- rbind2(fcs_raw_list[[1]], fcs_raw_list[[2]])
tmp2 <- rbind2(tmp, fcs_raw_list[[3]])
tmp3 <- rbind2(tmp2, fcs_raw_list[[4]])
fcs_rawer <- rbind2(tmp3, fcs_raw_list[[5]])
rm(tmp,tmp2,tmp3,res,fcs,b)

summary(res_rawer) # 1.9 to 32 million events per sample
c(dim(fcs_rawer[[1]]),dim(res_rawer[[1]])) # Check 1st and 3rd number are the same 

# Rename NK1.1
fcs_rawer <- fsApply(fcs_rawer, function(f){
  if (f@description$`$P12S` %in% "NK1p1"){f@description$`$P12S` <- "NK1.1"}
  if (f@description$`$P7S` %in% "NK1p1"){f@description$`$P7S` <- "NK1.1"}
  f
  })

```


Select live cells and their labels
```{r}
rm(manual)
rm(ff)
rm(labels.manual)
manual<-list() 
labels.manual <- list() # manually gated cell identity 
fcs_raw <- fcs_rawer 

for (d in 1:length(fcs_rawer)){
  manual[[d]] <- rep("Unknown", nrow(fcs_rawer[[d]]))
    for (i in 1:length(gateNames)){
    manual[[d]][res_rawer[[d]][,gateNames[i]]]<-counts_gate_names_my[i]
    }
  manual[[d]] <- factor(manual[[d]], levels = c(counts_gate_names_my,"Unknown"), ordered=TRUE)
  selected = res_rawer[[d]][,"live"] # TRUE FALSE vector for which cells you want to include (we only want to include live)
  fcs_raw[[d]]@exprs = exprs(fcs_rawer[[d]])[selected,]
}

newlevs <- c(counts_good_names_my[1],counts_good_names_my[2],counts_good_names_my[2],counts_good_names_my[1], counts_good_names_my[3:7], "Unknown") 
cbind(levels(manual[[1]][selected]), newlevs) # Check

for (d in 1:length(res_rawer)){
  selected = res_rawer[[d]][,"live"] 
  tmp <- manual[[d]][selected]
  levels(tmp) <- newlevs
  labels.manual[[d]] <- tmp
}


```

```{r}

fcs_empty_my <- fsApply(fcs_rawer, function(f){
  expr <- exprs(f)
  expr <- expr[1:2,]
  exprs(f) <- expr
  f
})
save(file="fsc_empty_my.R",fcs_empty_my)

```

flowAI to QC the data
```{r}
library(flowAI)

fcs_raw_qcing <- flow_auto_qc(fcs_raw, remove_from = "FS_FM", output = 2, timeCh = NULL,
      second_fractionFR = 0.1, alphaFR = 0.01, decompFR = TRUE,
      ChRemoveFS = c("FSC", "SSC"), outlierFS = FALSE, pen_valueFS = 200,
      max_cptFS = 3, ChFM = NULL, sideFM = "both", neg_valuesFM = 1,
      html_report = "_QC", mini_report = "QCmini", fcs_QC = FALSE,
      fcs_highQ = FALSE, fcs_lowQ = FALSE, folder_results = "resultsQC_post_my") # Removing signal acquisition (FS) and dynamic range (FM) abnormalities

# Now do again to get the actual flowset to use (just change output to 1) 
 fcs_raw_qced <- flow_auto_qc(fcs_raw, remove_from = "FS_FM", output = 1, timeCh = NULL,
      second_fractionFR = 0.1, alphaFR = 0.01, decompFR = TRUE,
      ChRemoveFS = c("FSC", "SSC"), outlierFS = FALSE, pen_valueFS = 200,
      max_cptFS = 3, ChFM = NULL, sideFM = "both", neg_valuesFM = 1,
      html_report = "_QC", mini_report = "QCmini", fcs_QC = FALSE,
      fcs_highQ = FALSE, fcs_lowQ = FALSE, folder_results = "resQC_post_my") 

# Now need to apply this to the labels to get the indices only for the cells which passed QC
# First check dimensions are okay:
cbind(dim(fcs_raw_qced[[2]])[1],dim(fcs_raw_qcing[[2]])[1],length(labels.manual[[2]])) # Last two should be the same
 
labels.manual.qced <- NULL
for (i in 1:length(labels.manual)){
labels.manual.qced[[i]] <- labels.manual[[i]][fcs_raw_qcing[[i]]@exprs[,"QCvector"] < 10000] 
}
labels.manual.qced.short.my <- lapply(labels.manual.qced, head, 50000)

fcs_raw_qced_short_my <- NULL
for (i in 1:length(fcs_raw_qced)){
  fcs_raw_qced_short_my[[i]] <- head(fcs_raw_qced[[i]], 50000)
} 

min(sapply(labels.manual.qced, length)) 
tail(fcs_raw_qced_short_my[[40]])
tail(labels.manual.qced.short.my[[40]],30)
```


And save:
```{r}
save(file="fcs_raw_qced_my.Rmd",fcs_raw_qced_short_my)
save(file="labels.manual.qced_my.Rmd",labels.manual.qced.short.my)
```

