---
title: "Figure 3"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Meningeal differential count analysis. Makes:
  - Figure 3B
  - Figure 3C
  - Figure 5D
  - Figure S9F

```{r}
library(ggplot2)
library(tidyverse)
library(reshape2)
library(readxl)
library(corrplot)
library(cowplot)
library(DESeq2) 
library(lmtest)
library(here)

# get rid of scientific notation
options(scipen=999)

load(here("res/data_bred.R")) # Load the data frame containing counts for each mouse
```

```{r}
# Select meningeal counts
tmp <- colnames(data_bred)[grep("men_",colnames(data_bred))]

# Define cell types
countdata <- c("men_MHCII+","men_MHCII-","men_helper","men_cytotoxic","men_microglia","men_neut","men_nk","men_nkt","men_b_both","men_plasmacytoid","men_cd11b+","men_cd11b-","men_ly6c.mono")

# Label cell types
counts_good_names_men <- c("CD64^{'+'}*MHCII^{'+'}~macrophages","CD64^{'+'}*MHCII^{'-'}~macrophages","T[helper]","T[cytotoxic]","microglia","neutrophils","NK","NKT","B~cells","pDCs","CD11b^{'+'}~DCs","CD11b^{'-'}~DCs","Ly6C^{hi}~monocytes") 

# Mean counts in WT mice
tmp <- data.frame(t(rbind(counts_good_names_men,signif(colMeans(data_bred[data_bred$strain=="wt" & data_bred$cond=="hc",countdata], na.rm=TRUE),2))))
rownames(tmp) <- tmp$counts_good_names_men
print(tmp)

```

Pie chart (Figure 3B inset)
```{r}

blank_theme <- theme_minimal() +
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.ticks = element_blank(),
  plot.title = element_text(size=14, face="bold")
  )

# Colors
colors_ord = c( 
  "#DC050C", 
  "#E6AB02", 
  "#882E72", 
  "#B17BA6", 
  "#55A1B1",
  "#8DD3C7", 
  "#33A02C", 
  "#8600bf",
  "#aeae5c", 
  "#FF7F00", 
  "#d4b7b7", 
  "#1965B0",
  "#ffff00", 
  "grey79" 
)

# Uncategorised cells
data_bred$unknown_men <- data_bred$men_leuco_plus_micro - rowSums(data_bred[,countdata])

# Order cells by median count
wt_counts_ordered <- sort(apply(data_bred[data_bred$strain=="wt" & data_bred$cond=="hc",countdata],2,median), decreasing=TRUE) 

names_ordered <- counts_good_names_men[order(apply(data_bred[data_bred$strain=="wt" & data_bred$cond=="hc",countdata],2,median), decreasing=TRUE)]

df <- data.frame(
  Cell = c(names_ordered, "Unknown"),
  value = c(wt_counts_ordered, median(data_bred[data_bred$strain=="wt" & data_bred$cond=="hc","unknown_men"], na.rm=TRUE))
)

df$Cell <- factor(df$Cell, levels=df$Cell, ordered=TRUE)

p <- ggplot(df, aes(x = "", y = value, fill = Cell)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = pi / 3) +
  scale_fill_manual(values = colors_ord, labels=parse(text=levels(df$Cell))) +
  blank_theme +
  theme(axis.text=element_blank(), legend.text=element_text(size=14), legend.text.align = 0, legend.title=element_blank()) 
p
ggsave(p, filename="apollo_meninges_counts_wt_pie.pdf", path=here("pics/"), width=6, height=5)
```

Percentage plot (Figure 3B)
```{r}
# Column names for percentages are same as for absolute counts but start with "pen" rather than "men"
percdata <- countdata
substring(percdata,1,1) <- "p"

percing <- data_bred[data_bred$strain %in% "wt" & data_bred$cond %in% "hc",c("name",percdata)]
percing$unknown_pen <- 1 - rowSums(percing[,percdata])
colnames(percing) <- c("name",counts_good_names_men,"Unknown")
percing_melt <- melt(percing)

percing_melt$variable <- factor(percing_melt$variable, levels=levels(df$Cell))

p <- ggplot(percing_melt, aes(x=variable, y=100*value, fill=variable)) + geom_boxplot(outlier.alpha=0) + geom_jitter(width=0.2, height=0, size=1, shape=16) + ylab("% of meningeal \nimmune cells") + xlab("") + guides(fill=FALSE) + theme_classic() + scale_x_discrete(limits=rev(levels(df$Cell)),labels=parse(text=rev(levels(df$Cell)))) + theme(panel.grid = element_blank()) + scale_fill_manual(values=colors_ord) + coord_flip()
p
ggsave(filename="apollo_percentages_hc_wt_meninges.pdf",p,path=here("pics/"),width=6,height=4)
# 13 mice

```

In the above, `r nrow(percing)` HC, non-stressed mice went into making up the plots.

## DEseq
```{r}
count_men <- data_bred[,countdata]
subs <- rownames(na.omit(count_men))
counts <- t(na.omit(as.matrix(count_men)))
colnames(counts) <- data_bred[subs,"name"]
condition <- relevel(factor(data_bred[subs,"cond"], ordered=FALSE), ref="hc") # DEseq needs ordering removed
strain <- relevel(factor(data_bred[subs,"strain"], ordered=FALSE), ref="wt")
cohort <- factor(data_bred[subs,"coh"],ordered=FALSE)

# Create a data frame of the design matrix as required by DESeqDataSetFromMatrix
coldata <- data.frame(row.names=colnames(counts), condition, strain, cohort) 

dds <- DESeqDataSetFromMatrix(countData=counts, colData=coldata, design= ~ cohort + strain + condition + strain:condition) 
dds

# DESeq including estimation of size factors
dds_libcorr <- DESeq(dds)

```


## DESeq without library size correction (sensitivity analysis)
```{r}
dds_nocorr <- dds
sizeFactors(dds_nocorr) <- sizeFactors(dds_libcorr)
sizeFactors(dds_nocorr) <- 1 # Make all size factors = 1
dds_nocorr <- estimateDispersions(dds_nocorr)
dds_nocorr <- nbinomWaldTest(dds_nocorr)

```


Might the library size correction introduce bias? If there is a significant effect of condition or strain on the calculated size factors (i.e. library size correction), this would suggest yes, but there is not:
```{r}
SF <- sizeFactors(dds_libcorr)
hist(SF)
data_bred$men_size_facs <- NULL
data_bred <- merge(data_bred, data.frame(name=names(SF),men_size_facs=SF), by="name")
ggplot(aes(y=men_size_facs, x=men_leucocytes, color=strain, shape=cond), data=data_bred) + geom_point() + scale_color_manual(values=c("red", "black"))

# Linear model of size factors against experimental design
rs <- lm(data_bred$men_size_facs ~ strain + cond + coh, data=data_bred, na.action=na.omit)
summary(rs) # See effects of cohort as expected, but no effects of strain or condition on size factors

# Coefficients:
#              Estimate Std. Error t value   Pr(>|t|)    
# (Intercept)  0.544863   0.132433   4.114   0.000168 ***
# straincd19   0.166743   0.096551   1.727   0.091183 .  
# cond.L      -0.009125   0.069272  -0.132   0.895807    
# cohr4        0.690711   0.174402   3.960   0.000270 ***
# cohr5        0.833322   0.174402   4.778 0.00001999 ***
# cohr6        0.999690   0.176598   5.661 0.00000106 ***
# cohr7        0.806056   0.174402   4.622 0.00003325 ***
# cohr8        0.510212   0.174402   2.925   0.005421 ** 
# cohr9       -0.378343   0.199344  -1.898   0.064277 . 

```

## Save rlog counts
```{r}
dds <- dds_libcorr

rld <- rlogTransformation(dds, blind=TRUE) 
men_rld_libcorr <- assay(rld) 

save(men_rld_libcorr, file=here("res/men_rld_libcorr.R"))
save(counts_good_names_men, file=here("res/counts_good_names_men.R"))
```

### Effects of stress in WT mice
```{r}
library(ggrepel)
fdr <- 0.05
res <- results(dds,name="condition_csd_vs_hc")
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=TRUE)), by="row.names", sort=FALSE) 
names(resdata)[1] <- "Subset"
head(resdata[order(resdata$padj),c("Subset","pvalue","padj")]) # Look at min p values (to decide y-axis limits)
head(resdata[order(abs(resdata$log2FoldChange), decreasing=TRUE),c("Subset","log2FoldChange")]) # Look at max fold change (to decide x-axis limits)

res2 = mutate(resdata, sig=ifelse(resdata$padj<fdr, paste("FDR<",fdr,sep=""), "Not Sig"))
rownames(res2) <- counts_good_names_men

p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme_bw() +
  theme(legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="grey", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) + 
  xlim(-1.5,1.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p

p2 = p + geom_text_repel(data=dplyr::filter(res2, res2$padj<fdr), aes(dplyr::filter(res2, res2$padj<fdr)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<fdr)[,"pvalue"]),label=rownames(res2[res2$padj<fdr,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) 
p2
ggsave(p2,path=here("pics/"), filename="apollo_volcano_counts_stress_hc_libcorr.pdf",width=6,height=6)

# Version with more labels
fdrT <- 0.5
p3 = p + geom_text_repel(data=dplyr::filter(res2, res2$padj<fdrT), aes(dplyr::filter(res2, res2$padj<fdrT)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<fdrT)[,"pvalue"]),label=rownames(res2[res2$padj<fdrT,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) 
p3
ggsave(p3,path=here("pics/"), filename="apollo_volcano_counts_stress_hc_libcorr_moretext.pdf",width=6,height=6)
res2_men_stress_hc_libcorr <- res2

```

### Effect of strain 
Figure 5D
```{r}
res <- results(dds,name="strain_cd19_vs_wt")
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=TRUE)), by="row.names", sort=FALSE)
names(resdata)[1] <- "Subset"
head(resdata)

# Had to alter the x limits manually
res2 = mutate(resdata, sig=ifelse(resdata$padj<fdr, paste("FDR<",fdr,sep=""), "Not Sig"))
rownames(res2) <- counts_good_names_men

p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme_bw() +
  theme(legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="grey", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) + 
  xlim(-1.5,1.5) +
  #ylim(0,3.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p

p = p + geom_text_repel(data=dplyr::filter(res2, res2$padj<fdr), aes(dplyr::filter(res2, res2$padj<fdr)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<fdr)[,"pvalue"]),label=rownames(res2[res2$padj<fdr,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) 
p
ggsave(p,path=here("pics/"), filename="apollo_volcano_counts_cd19_wt_libcorr.pdf",width=6,height=6)
res2_men_strain_wt_libcorr <- res2

```

### EFFECT OF STRESS IN CD19 MICE:
Figure S9F
```{r}
res <- results(dds,list(c("condition_csd_vs_hc","straincd19.conditioncsd")))
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata)[1] <- "Subset"
head(resdata)

# Had to alter the x limits manually

res2 = mutate(resdata, sig=ifelse(resdata$padj<fdr, paste("FDR<",fdr,sep=""), "Not Sig"))
rownames(res2) <- counts_good_names_men

p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme_bw() +
  theme(legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="grey", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) + 
  xlim(-1.5,1.5) +
  #ylim(0,3.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p

# Add labels
p = p + geom_text_repel(data=dplyr::filter(res2, res2$padj<fdr), aes(dplyr::filter(res2, res2$padj<fdr)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<fdr)[,"pvalue"]),label=rownames(res2[res2$padj<fdr,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) 
p
ggsave(p,path=here("pics/"), filename="apollo_volcano_counts_stress_in_cd19_libcorr.pdf",width=6,height=6)

res2_men_stress_cd19_libcorr <- res2
save(file=here("/res/res2_men.R"), res2_men_stress_hc_libcorr, res2_men_strain_wt_libcorr, res2_men_stress_cd19_libcorr)

```

Double-check names
```{r}
print(rbind(rownames(res2),res2$Subset))
print(summary(rs))
```

## DESeq without library size correction (sensitivity analysis)
```{r}
dds <- dds_nocorr 
resultsNames(dds)
```

### Effects of stress
```{r}
res <- results(dds,name="condition_csd_vs_hc")
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=FALSE)), by="row.names", sort=FALSE) # Normalized = TRUE or FALSE doesn't make a difference here because Log fold change already calculated by this stage anyway.
names(resdata)[1] <- "Subset"
head(resdata[order(resdata$padj),])

res2 = mutate(resdata, sig=ifelse(resdata$padj<0.05, "FDR<0.05", "Not Sig"))
rownames(res2) <- counts_good_names_men
library(ggrepel)
p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  geom_text_repel(data=dplyr::filter(res2, res2$padj<0.05), aes(dplyr::filter(res2, res2$padj<0.05)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<0.05)[,"pvalue"]),label=rownames(res2[res2$padj<0.05,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme(panel.background = element_rect(colour="white"), legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="black", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) +
  xlim(-1.5,1.5) +
  #ylim(0,3.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p
ggsave(p,path=here("pics/"), filename="apollo_volcano_counts_stress_hc_nocorr.pdf",width=6,height=6)

```


### Effect of strain (raw, unnormalized counts)
```{r}
res <- results(dds,name="strain_cd19_vs_wt")
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata)[1] <- "Subset"
head(resdata)

res2 = mutate(resdata, sig=ifelse(resdata$padj<0.05, "FDR<0.05", "Not Sig"))
rownames(res2) <- counts_good_names_men

p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  geom_text_repel(data=dplyr::filter(res2, res2$padj<0.05), aes(dplyr::filter(res2, res2$padj<0.05)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<0.05)[,"pvalue"]),label=rownames(res2[res2$padj<0.05,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme(panel.background = element_rect(colour="white"),legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="black", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) + 
  xlim(-1.5,1.5) +
  #ylim(0,3.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p
ggsave(p,path=here("pics/"), filename="apollo_volcano_counts_cd19_wt_nocorr.pdf",width=6,height=6)

```

### EFFECT OF STRESS IN CD19 MICE:
```{r}
res <- results(dds,list(c("condition_csd_vs_hc","straincd19.conditioncsd")))
resdata <- merge(as.data.frame(res), as.data.frame(counts(dds, normalized=FALSE)), by="row.names", sort=FALSE)
names(resdata)[1] <- "Subset"
head(resdata)

res2 = mutate(resdata, sig=ifelse(resdata$padj<0.05, "FDR<0.05", "Not Sig"))
rownames(res2) <- counts_good_names_men

p = ggplot(res2, aes(res2$log2FoldChange, -log10(res2$pvalue))) +
  geom_point(aes(col=sig, size=sig)) +
  scale_color_manual(values=c("red", "black")) +
  scale_size_manual(values=c(3,2)) +
  geom_text_repel(data=dplyr::filter(res2, res2$padj<0.05), aes(dplyr::filter(res2, res2$padj<0.05)[,"log2FoldChange"], -log10(dplyr::filter(res2, res2$padj<0.05)[,"pvalue"]),label=rownames(res2[res2$padj<0.05,])),min.segment.length = unit(0, "lines"), size=7,segment.colour = "black", parse=TRUE, box.padding=unit(0.4,"lines")) +
  xlab("log2(Fold Change)") +
  ylab("-log10 (p-value)") +
  theme(panel.background = element_rect(colour="white"), legend.justification=c(0,0), legend.position=c(0,0), legend.background = element_rect(color="black", fill="white",size=.5, linetype="solid"), legend.text=element_text(size=20), axis.text = element_text(size=20), axis.title = element_text(size=20)) + 
  xlim(-1.5,1.5) +
  #ylim(0,3.5) +
  labs(col=NULL) +
  guides(size=FALSE, col=guide_legend(keywidth=0.05,keyheight=1.5)) +
  geom_vline(xintercept=0, linetype="dashed")
p
ggsave(p,path=here("pics/"), filename="apollo_volcano_counts_stress_in_cd19_no_corr.pdf",width=6,height=6)

```


```{r}

sessionInfo()

```


